---
title: "Dada2_report"
author: "Ramón Gallego"
date: "1/9/2018"
output: html_document
params:
  folder: 
    value: ../data_sub/noprimers
  hash: Yes
  trimming.length.Read1: 200
  trimming.length.Read2: 200
  metadata: output.metadata.csv
  output.folder: outputs

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=params$folder)
```

## Dada2 report

You have successfully split your libraries into a pair (or two) of fastq files per sample. Now let's import the output of demultiplex_both_fastq.sh

First load the packages. And let's update the parameters we need
```{r loading packages, echo=FALSE ,message=FALSE}

library (tidyverse)
library (dada2)
library (digest)
library (seqinr)
library (knitr)
library (kableExtra)

sample.metadata <- read_csv(params$metadata)
filt_path <- file.path(params$folder, "filtered")
getwd()
filt_path
# Check if output directory exists - if not create as a subfolder of input dir

if(!dir.exists(file.path(params$output.folder))){
  dir.create(file.path(params$folder, params$output.folder))
  output.dir <- file.path(params$folder, params$output.folder)
}else{
  output.dir <- file.path(params$output.folder)
}

```

```{r check quality trim point, message=FALSE, warning=FALSE}


sample.metadata %>% 
  sample_n(4) %>% pull(file1) %>% 
  plotQualityProfile(.)

sample.metadata %>% 
  sample_n(4) %>% pull(file2) %>% 
  plotQualityProfile(.)

```

The most common Amplicon length is 302, so trimming to 200 bp each read should give us enough overlap to successfully merge Forward and Reverse reads. Based on the drop of quality from the reverse read, you might choose to make them shorter on the R2. In that case, rerun the script setting the trimming differently.

```{r dadaing}

sample.metadata %>% 
  mutate(filtF1 = file.path("filtered", str_replace(file1, "_L001_R1_001.fastq_sub.fastq", "_F1_filt.fastq.gz")),
         filtR1 = file.path("filtered", str_replace(file2, "_L001_R2_001.fastq_sub.fastq", "_R1_filt.fastq.gz"))) %>% 
  mutate (outFs = pmap(.l= list (file1, filtF1, file2, filtR1),
                       .f = function(a, b, c, d) {
                         filterAndTrim(a,b,c,d,
                                       truncLen=c(params$trimming.length.Read1,params$trimming.length.Read2),
                                       maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                                       compress=TRUE, multithread=TRUE )
                       } ),
          errF1 = map(filtF1, ~ learnErrors(.x, multithread=TRUE,verbose = 0)),     # Calculate errors
          errR1 = map(filtR1, ~ learnErrors(.x, multithread=TRUE,verbose = 0)),
          derepF1 = map(filtF1, derepFastq),                   # dereplicate seqs
          derepR1 = map(filtR1, derepFastq),
          dadaF1  = map2(derepF1,errF1, ~ dada(.x, err = .y, multithread = TRUE)),  # dada2
          dadaR1  = map2(derepR1,errR1, ~ dada(.x, err = .y, multithread = TRUE)),
          mergers = pmap(.l = list(dadaF1,derepF1, dadaR1,derepR1),                 # merge things
                         .f = mergePairs )) -> output.dada2

write_rds(output.dada2, path = "output.halfway.rds")
seqtabF <- makeSequenceTable(output.dada2$mergers)
dim(seqtabF)


table(nchar(getSequences(seqtabF)))

```
Firstly, we'll find the patterns that separate our Fwd, Rev, .1 and .2 files. look at the quality of the .1 and .2 reads
Run dada 2 step by step
```{r RemovingChimeras, message=F}
seqtab.nochim <- removeBimeraDenovo(seqtabF, method="consensus", multithread=TRUE)
dim(seqtab.nochim)
seqtab.nochim.df <- as.data.frame(seqtab.nochim)
```


```{r}
# Output files
  conv_file <- file.path(output.dir,"hash_key.csv")
  conv_file.fasta <- file.path(output.dir,"hash_key.fasta")
  
  ASV_file <-  file.path(output.dir,"ASV_table.csv")
# Now decide if you want hashing or not
if ( params$hash==TRUE)
  {
  
  conv_table <- tibble( Hash = "", Sequence ="")
  
  hashes <- list(NULL)
 
######## Create the Hash key file
  
  for (i in 1:ncol(seqtab.nochim.df)) {   #for each column of the dataframe
    current_seq <-colnames(seqtab.nochim.df)[i] # Take the whole sequence from each column of the dataframe
    current_hash <- digest(current_seq,algo = "sha1",serialize = F,skip = "auto") # Hash it so it is both shorter and unique
    hashes[[i]] = current_hash
    conv_table [i,]<-c(current_hash, current_seq) # add the Hash - sequence conversion to a table
    colnames(seqtab.nochim.df)[i] <- current_hash
  }
  write_csv(conv_table, conv_file) # write the table into a file
  
  write.fasta(sequences = as.list(conv_table$Sequence),
              names     = as.list(conv_table$Hash),
              file.out = conv_file.fasta)

 seqtab.nochim.df <- bind_cols(sample.metadata %>% 
                                select(Sample_name, Locus),
                              seqtab.nochim.df)


seqtab.nochim.df %>% 
  pivot_longer(cols = c(- Sample_name, - Locus),
              names_to = "Hash",
              values_to = "nReads") %>% 
  filter(nReads > 0) -> current_asv
write_csv(current_asv, ASV_file)    }else{
  #What do we do if you don't want hashes: two things - Change the header of the ASV table, write only one file

   seqtab.nochim.df %>% 
  pivot_longer(cols = c(- Sample_name, - Locus),
              names_to = "Sequence",
              values_to = "nReads") %>% 
  filter(nReads > 0) -> current_asv
write_csv(current_asv, ASV_file) 
  
}
```


##Track the fate of all reads
```{r output_summary}

getN <- function(x) sum(getUniques(x))

output.dada2 %>% 
  select(-file1, -file2, -filtF1, -filtR1, -errF1, -errR1, -derepF1, -derepR1) %>% 
  mutate_at(.vars = c("dadaF1", "dadaR1", "mergers"),
            ~ sapply(.x,getN)) %>% 
#  pull(outFs) -> test
  mutate(input = map_dbl(outFs, ~ .x[1]),
         filtered = map_dbl(outFs, ~ .x[2]),
         tabled  = rowSums(seqtabF),
         nonchim = rowSums(seqtab.nochim)) %>% 
  select(Sample_name,
         Locus,
         input,
         filtered,
         denoised_F = dadaF1,
         denoised_R = dadaR1,
         merged = mergers,
         tabled,
         nonchim) -> track

write_csv(track, file.path(params$output.folder,"dada2_summary.csv"))

kable (track, align= "c", format = "html") %>%
      kable_styling(bootstrap_options= "striped", full_width = T, position = "center") %>%
      column_spec(2, bold=T) 
track %>% 
  mutate_if(is.numeric, as.integer) %>% 
  pivot_longer(cols = c(-Sample_name, -Locus),
               names_to = "Step",
               values_to = "Number of Sequences") %>% 
  mutate (Step = fct_relevel(Step, 
                             levels = c( "input","filtered","denoised_F" ,"denoised_R" , "merged" , "tabled", "nonchim"))) %>% 
  ggplot(aes(x = Step, y = `Number of Sequences`, group =  Sample_name, color = Sample_name)) +
  geom_line() +
  facet_wrap(~Locus) +
  guides(color = "none")

```
