---
title: "Create Illumina Sample Sheet and metadata"
author: "Ramon Gallego"
date: "11/8/2019"
output: html_document
params: 
  Assay: "Nextera XT"
  Index_Adapters: "Nextera XT Index Kit (96 Indexes 384 Samples)"
  Cycles_per_pairend: 301
  Set: 1
  input.metadata: ../data_sub/metadata.csv
  date: !r Sys.Date()a
  output_dir: ../data_sub/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## What does this do

It creates the Illumina Sample sheet and the `metadata.csv` file that you need for running the pipeline. You need to double check the output - weird things happen all the time.

## Required input

You need an input file with the following columns:

  - Sample name
  - Well in the 96-well plate used for the Nextera reaction. 
  - PrimerF, 
  - PrimerR,
  - Plate, refers to the set of of nextera adapters used.


## Loading the metadata you want to sequence

```{r load metadata}

init.metadata <- read_csv(params$input.metadata)

init.metadata

```

### Check that the metadata has all needed fields

```{r Check metadata}

init.metadata %>% 
  rownames_to_column("Sample_number") %>% 
  mutate_all(as.character) %>% 
  pivot_longer(-Sample_number, names_to = "Variable", values_to = "Value") %>% 
  summarise(Sample_name = case_when(sum(str_detect(Variable, "Sample")) > 0 ~ "Present",
                                 TRUE                                    ~ "Absent"),
         Well        = case_when(sum(str_detect(Variable, "Well")) > 0 ~ "Present",
                                 TRUE                                    ~ "Absent"),
         Primer_F    = case_when(sum(str_detect(Variable, "Primer_F")) > 0 ~ "Present",
                                 TRUE                                    ~ "Absent"),
         Primer_R    = case_when(sum(str_detect(Variable, "Primer_R")) > 0 ~ "Present",
                                 TRUE                                    ~ "Absent"),
         
         Plate       = case_when(sum(str_detect(Variable, "Plate")) > 0 ~ "Present",
                                 TRUE                                    ~ "Absent")) -> Checks
  




```


## Load the Illumina SampleSheet template



```{r Loading SampleSheet}
template.sample.sheet <- read_lines("../data_sub/SampleSheetLibPrep.csv")
```

Locate all the lines in which the new parameters are going to be written

```{r locate terms}


date.row  <- str_which(template.sample.sheet, "^\\Date")
Assay.row <- str_which(template.sample.sheet, "^Assay")
Index.row <- str_which(template.sample.sheet, "^Index Adapters")
Reads.row <- str_which(template.sample.sheet, "^\\[Reads]")
data.start <- str_which(template.sample.sheet, "^\\[Data]")


```

## Load the Illumina adapters

```{r load nextera}

Nextera_Adapters_i7 <- read_csv("../data_sub/Nextera_adapters_i7.csv")

Nextera_Adapters_i5 <- read_csv("../data_sub/Nextera_adapters_i5.csv")

Nextera_Adapters_i7 %>% 
  mutate (Set = rep(c(1,2,3), each = 12)[1:26]) -> Nextera_Adapters_i7

Nextera_Adapters_i5 %>% 
  mutate (Set = rep(c(1,2,3), each = 8)[1:18]) -> Nextera_Adapters_i5

```

## Merge metadata with illumina adapters

```{r merging}
Nextera_Adapters_i7 %>% 
  filter (Set == params$Set) %>% 
  select(-Position, -`Bases in Adapter`, -Set) %>% 
  right_join(init.metadata, by = c("Column")) %>% 
  left_join(Nextera_Adapters_i5 %>% 
              filter(Set == params$Set) %>% 
              select(-Position, -`Bases in Adapter`)) %>% 
  select(Sample, everything()) -> metadata


```

## Fill Illumina SampleSheet

First with the parameters not linked with each sample

```{r}
template.sample.sheet[date.row] <- paste0("Date,",Sys.Date())

template.sample.sheet[Assay.row] <- paste0("Assay," ,params$Assay)
template.sample.sheet[Index.row] <- paste0("Index Adapters,",params$Index_Adapters)
template.sample.sheet[Reads.row +2 ] <- template.sample.sheet[Reads.row + 4] <- params$Cycles_per_pairend
```

Then with the rest of the dataset

```{r}
sample.data <- tibble (Sample_Name = metadata$Sample,
                      Sample_Plate = metadata$Set,
                      Sample_Well = metadata$Well,
                      I7_Index_ID = metadata$I7_Index_ID,
                      index       = metadata$index,
                      I5_Index_ID = metadata$I5_Index_ID,
                      index2      = metadata$index2,
                      Sample_Project ="", 
                      Description = "") %>% 
                rownames_to_column("Sample_ID")


nsamples <- nrow(sample.data)
template.sample.sheet <- template.sample.sheet[1:data.start]
for (i in 1:nsamples){
  
  template.sample.sheet[i+data.start] <- paste(sample.data[i,], collapse = ",") 
  
  }
template.sample.sheet

write_lines(template.sample.sheet,  paste0(params$output_dir, "SampleSheet_",Sys.Date(),".csv"))
```

## Fill in metadata

We are going to assume that the Illumina always returns files with the format `Sample_name` `_L001_R` [1-2] '_001.fastq`. So let's fill the metadata accordingly

```{r}
metadata %>% 
  mutate(file1 = paste0(Sample, "_L001_R1_001.fastq"),
         file2 = paste0(Sample, "_L001_R2_001.fastq")) %>% 
  select(Sample,
         file1,
         file2,
         I7_Index_ID,
         I5_Index_ID,
         Well,
         Set,
         Primer_F,
         Primer_R) -> metadata


write_csv(metadata, paste0(params$output_dir, "metadata_",Sys.Date(),".csv"))
```

